The 2- and 3-point surface correlation functions calculations: from novel exact continuous approach to improving methodology for discrete images

### 2023-05-17 paper.tex
Replace noise -> perturbation

In majority, if not all, cases mentioned above, surface functions - namely,
$F_{ss}$ and $F_{sv}$ - were underrepresented due to our inability to compute
them in a robust way. While known for a long time \cite{dietrich1995scattering}, only
somewhat crude approximate approaches \cite{seaton1986spatial} were developed
until Ma and Torquato laid foundation for such a robust computational
methodology \cite{ma2018SS}. They proposed an elegant solution for infinite
resolution Gaussian random fields, or "continuous" approach. They also assumed
that the same approach can be readily applied to discrete experimental images if
grey-scale values are thresholded similarly to continuous Gaussian
field. Samarin et al. \cite{Samarin} argued that experimental images needs
to be segmented first using current cutting edge techniques \cite{NNseg} and
proposed "discrete" approach. The latter allowed obtaining surface functions
closer to analytical solutions for overlapping disks/spheres and to suppress
noise, in particular for $F_{ss}$. The noise in Ma and Torquato method resulted
from $1/cos(\theta)$ computations during random line segment placement - this
leaves a room for improvements. While Samarin et al. obtained better convergence
with analytical solution, this was demonstrated only for overlapping disks image
of limited size (that can potentially introduce some differences between
computations and analytical solution for infinite size structure). Moreover,
good agreement was achieved by manually fitting the values of edge-detecting
filter that could lack accuracy for arbitrary structure different from
overlapping disks. In other words, we still lack robustness in surface CFs
evaluation due to inability to directly compare "discrete" versus "continuous"
methodology, and verify both approaches for arbitrary geometries.

@@

In this contribution we continue to build upon foundational work of Ma and
Torquato \cite{ma2018SS} and recent findings of Samarin et al. \cite{Samarin} to
develop a robust and efficient approach to compute surface correlation functions
from digital 2D and 3D images. In particular, here our aim is to establish a
direct verification of "discrete" methodology (that is easily parallelizable on
CPU and GPU architectures) using exact noiseless "continuous" approach for
arbitrary structures (still described by some continuous function). The rest of
the manuscript is organized as follows: in \cref{sec:def} we introduce some
major definitions of CFs necessary for establishing algorithms, in
\cref{sec:algo} we describe the all methodological details for "discrete" method
for digital images. Next, in \cref{sec:algo-precise} we explain in detail why
verification of computational approaches for digital images is currently
hampered, and then introduce novel exact "continuous" methodology in
\cref{sec:fss-2d}. The choice of verification cases is presented in
\cref{sec:results} followed by improvements in methodology of Samarin et~al. The
paper concludes with discussion, outline and summary in \cref{sec:summary}.

@@

\section{Discussion, outlook and summary}
\label{sec:summary}
The exact continuous method presented in this paper is developed only for
surface-surface function (and its 3-point counterpart). The surface-void
function computations in digital manner are much more robust and exhibit no
noise \cite{ma2018SS,Samarin}. The filter $H'$ verified for $F_{ss}$ also
improves calculations for $F_{sv}$ (as evident from \cref{fig:not-covered}).

>>>>

In majority, if not all, cases mentioned above, surface functions - namely,
$F_{ss}$ and $F_{sv}$ - were underrepresented due to our inability to compute
them in a robust way. While known for a long time \cite{dietrich1995scattering}, only
somewhat crude approximate approaches \cite{seaton1986spatial} were developed
until Ma and Torquato laid foundation for such a robust computational
methodology \cite{ma2018SS}. They proposed an elegant solution for infinite
resolution Gaussian random fields, or "continuous" approach. They also assumed
that the same approach can be readily applied to discrete experimental images if
grey-scale values are thresholded similarly to continuous Gaussian
field. Samarin et al. \cite{Samarin} argued that experimental images needs
to be segmented first using current cutting edge techniques \cite{NNseg} and
proposed "discrete" approach. The latter allowed obtaining surface functions
closer to analytical solutions for overlapping disks/spheres and to suppress
perturbation, in particular for $F_{ss}$. The perturbation in Ma and Torquato
method resulted from $1/cos(\theta)$ computations during random line segment
placement -- this leaves a room for improvements. While Samarin et al. obtained
better convergence with analytical solution, this was demonstrated only for
overlapping disks image of limited size (that can potentially introduce some
differences between computations and analytical solution for infinite size
structure). Moreover, good agreement was achieved by manually fitting the values
of edge-detecting filter that could lack accuracy for arbitrary structure
different from overlapping disks. In other words, we still lack robustness in
surface CFs evaluation due to inability to directly compare "discrete" versus
"continuous" methodology, and verify both approaches for arbitrary geometries.

@@

In this contribution we continue to build upon foundational work of Ma and
Torquato \cite{ma2018SS} and recent findings of Samarin et al. \cite{Samarin} to
develop a robust and efficient approach to compute surface correlation functions
from digital 2D and 3D images. In particular, here our aim is to establish a
direct verification of "discrete" methodology (that is easily parallelizable on
CPU and GPU architectures) using exact "continuous" approach for arbitrary
structures (still described by some continuous function). The rest of the
manuscript is organized as follows: in \cref{sec:def} we introduce some major
definitions of CFs necessary for establishing algorithms, in \cref{sec:algo} we
describe the all methodological details for "discrete" method for digital
images. Next, in \cref{sec:algo-precise} we explain in detail why verification
of computational approaches for digital images is currently hampered, and then
introduce novel exact "continuous" methodology in \cref{sec:fss-2d}. The choice
of verification cases is presented in \cref{sec:results} followed by
improvements in methodology of Samarin et~al. The paper concludes with
discussion, outline and summary in \cref{sec:summary}.

@@

\section{Discussion, outlook and summary}
\label{sec:summary}
The exact continuous method presented in this paper is developed only for
surface-surface function (and its 3-point counterpart). The surface-void
function computations in digital manner are much more robust and exhibit less
perturbation \cite{ma2018SS,Samarin}. The filter $H'$ verified for $F_{ss}$ also
improves calculations for $F_{sv}$ (as evident from \cref{fig:not-covered}).
